### 2019-05-23 thesis-template/msc-thesis.tex
update

% Choose language
\newcommand{\setlang}[1]{\selectlanguage{#1}\nonfrenchspacing}

@@

\section{Requirements and Ambitions}
\label{sec:spec} 
Our ambition is to develop a simple and efficient Global Illumination Propagation Scheme able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t. energy conservation and propagation attenuation principles. We then generate transmission profiles from the eigensystems (principal component frames) of the tensors for each cell in the domain (field). That is, to provide an orientation for a grid of crystal fibre structures (cnf. optical fibres) to redirect the intensities in analogy to the anisotropies of the tensor field within a glyph representation, obtained from Principal Component Analysis. Our approach should be implicitly adaptive to any kind of 2D dataset. Thus, we opt to be compatible to handle any kind of matrix (symm./anti-symm.) and any kind of resolution of the field (full-spectrum). At last, we aim to generate a new FTLE-related method for tensor field visualization by applying a Global Illumination Gradient to the resulting light distributions for every possible position and direction of the light source in the grid generated by stimulation with Delta-Impulses. Thus, our motivation is to segment key locations (LCS/ridges) in the field with tensor field lines (hyperstreamlines) converging/diverging, which is necessarily the same (for bidirectional tensor fields). Our approach should efficiently handle (relatively) large datasets. 
Knapp zwei Seiten, in dem die Anforderungen, die Zielsetzung und die
Methoden überblicksartig beschrieben werden. Hier sollte die
Beschreibung ``technischer'' bzw.~``formaler'' sein als in der
Einleitung, da der Leser nun mit den Grundlagen und verwandten
Arbeiten vertraut ist.

\section{Global Illumination Scheme}
\label{sec:scheme}
Within the scope of this work, we will develop a Global Illumination Scheme designed to efficiently propagate the intensities stored in discrete polar coordinates given as impulses (stimuli) by the user or the config. Polar profiles for different light sources are shown in Fig. \ref{polar}:
\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.8\textwidth]
    {img/strahler.png}
  }
  \caption{Types of forEmitters}
  \label{polar}
\end{figure}\vskip 3pt
Polar profiles map each angle $\varphi$ a magnitude $r$ forming a function $r(\varphi)$. They are circular axis versions of the domain $[0..2\pi]$, with negative magnitudes mirrored by 180 degrees). In the sense of interpretability, we exclude negative values (energies) from the scope of our calculations. As a start-up for the propagation, we will think of a grid cell in 2D as a center pixel within an 8-neighborhood. It has 8 (4 faces/edges and 4 diagonals) unique neighbors, which are adjacent point lights in the grid (the type of light source with theoretically no finite extend). We consider one face neighbor (top) to shortly explain the steps of the scheme. The diagonals and the others are obtained from derivation and symmetry.

@@

In diesem und den nachfolgenden Abschnitten werden die Beiträge der
Arbeit motiviert, formal sauber (oft mathematisch, sprich mit
Definitionen etc.) beschrieben, und bei Bedarf mithilfe von Beispielen
verdeutlicht. Die Beschreibungen in diesem Kapitel sind meist
unabhängig von einer konkreten Realisierung und Daten; diese werden im
nachfolgenden Kapitel detailliert.

\section{Transmission Profiles and Weighting Functions}
\label{sec:trans}
To obtain a kind of footprint of the tensor field, we set up ellipsoid equations for each glyph representation, obtained from applying PCA on each cell (tensor) in the field. We map the singular values (corresponding to eigenvalues) in decreasing order to the ellipses radii $a(\lambda_1 )$ and $b(\lambda_2 )$ as follows:

Ellipse Equation:
\begin{align}
	r(\omega) = \frac{ab}{\sqrt{a^{2}\sin^{2}(\omega-\varphi)+b^{2}\cos^{2}(\omega-\varphi)}} =\frac{\lambda_1\lambda_2}{\sqrt{\lambda_1^2\sin^2(\omega-\varphi)+\lambda_2^2\cos^2(\omega-\varphi)}}
\end{align}
That is, the singular values form the half-axes of the PCA ellipsoid, which yields us a symbolic definition for our transmission (transfer) function as weighting profiles. We compute the offset angle $\varphi$ w.r.t. x-axis by exploiting the atan2-function with 4-quadrant evaluation for the first singular vector. As a preliminary prep, this profile is precomputed (sampled) for all discrete angular steps for each cell (e.g. $360$). The weighting scheme which we apply is depicted in Fig. \ref{weighting}.
\begin{figure}[h]
  \centering
  {
    \includegraphics[width=0.6\textwidth]
    {img/weighting.png}
  }
  \caption{red: Intensity Profile, cyan: Transmission Profile (Glyph Ellipsoid Eq.)}
  \label{weighting}
\end{figure}
We interpret the polar functions as discrete, sampled 1D vectors with a constant angular resolution. That means, we use element-wise multiplication for the 1D-vectors in red (Intensity Profile) and cyan (Transmission Profile) to obtain a windowed version of the intensity profile and perform the following integration step to accumulate the total radiant power (flux) as follows. Note that we model angular dependent radiant intensity $\mathbf{I}(\omega) [\frac{W}{\mathop{rad}}]$ as an underlying physical entity, which is well suited for describing point light sources:

\begin{align}
	\Phi_t &= \int_0^{2\pi}\mathbf{I}(\omega)\mathbf{T}(\omega)\mathop{d\omega}\\
	\sum_{\cos} &= \int_0^{2\pi}\cos(\omega)\mathop{d\omega} \approx 2
\end{align}

This enables us to scale an oriented cosine-lobe w. the total radiant flux w. respect to its own total energy when integrated likewise. A discrete cosine lobe never reaches its theoretical sum of $2$ exactly in the interval $[0,2\pi]$. Therefore it needs to be computed to account for floating-point precision errors and normalized by following relation (e.g. $\sum_{\cos} = 1.9949204635834517$).
\begin{align}
	\cos_n(\omega) = \frac{\Phi_t}{\sum_{\cos}}cos(\omega)
\end{align}
This yields a cosine lobe, which has an overall energy of $\Phi_t$ if the point light is interpreted with a Lambert-Emitter (cnf. \cite{lindlein}) characteristic (as part or area element of a diffusely reflecting surface).

Now each tensor ellipse equation has a mean transmission rate $[\%]$, which turns out to be a measure or synonym for the total transmitted rate of the total incoming radiant intensity $\int_0^{2\pi}\mathbf{I}(\omega)$. Since energy can not be generated from transmission profiles (only transmitting, not generating), we have to normalize every tensor ellipse equation with the highest mean transmission rate in the field. This restricts the tensor ellipse mean to a maximum of 1.0 for the one with the highest tensor magnitude. We will use the magnitude $\bar{r}_{max}$ from the PCA here for simplicity. Hence the tensor ellipse equations are normalized through the following step:
\begin{align}
	r_n(\omega) = \frac{1}{\bar{r}_{max}}r(\omega)
\end{align}
Notice, that this implements the concept of absorption implicitly, because tensors with a lower mean $\bar{r}$ than $100\%$ lose energy wheras absorption equals $\boldsymbol{\mu} = \mathbf{1}-\mathbf{T}$ here. Now that we have coped with the problem of overall transmission, we are left with another problem that occurs with the weighting functions as that they can yield values above 1. This would potentially generate intensity out of nothing which leads to crucially unstable behaviour. Well, but only in case energy conservation principles were not respected. However, a proper normalization can give corrective remedy. To construct a solution, we ask for the source of the excessive energy. If the tensor mean complies with the constraint $\bar{r}\leq 1.0$, then the energy must lack in some other place (since the overall transmission fulfills the energy conservation constraint in that case). 
\section{Global Illumination Gradient}
\label{sec:ftle}
Note that the resulting scalar field is $(n+1)$-dimensional ($x,y,z$ and direction $\theta$) and needs to be flattened by averaging or projection in 3D for proper visualization. 
Usw.

>>>>

% Choose language
\newcommand{\setlang}[1]{\selectlanguage{#1}\nonfrenchspacing}
\newcommand*\mean[1]{\bar{#1}}

@@

\section{Requirements and Ambitions}
\label{sec:spec} 
Our ambition is to develop a simple and efficient Global Illumination Propagation Scheme able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t. energy conservation and propagation attenuation principles. We then generate transmission profiles from the eigensystems (principal component frames) of the tensors for each cell in the domain (field). That is, to provide an orientation for a grid of crystal fibre structures (cnf. optical fibres) to redirect the intensities in analogy to the anisotropies of the tensor field within a glyph representation, obtained from Principal Component Analysis. Our approach should be implicitly adaptive to any kind of 2D dataset. Thus, we opt to be compatible to handle any kind of matrix (symm./anti-symm.) and any kind of resolution of the field (full-spectrum). At last, we aim to generate a new FTLE-related method for tensor field visualization by applying a Global Illumination Gradient to the resulting light distributions for every possible position and direction of the light source in the grid generated by stimulation with Delta-Impulses. Thus, our motivation is to segment key locations (LCS/ridges) in the field with tensor field lines (hyperstreamlines) converging/diverging, which is necessarily the same (for bidirectional tensor fields). Our approach should efficiently handle (relatively) large datasets. 
%Knapp zwei Seiten, in dem die Anforderungen, die Zielsetzung und die
%Methoden überblicksartig beschrieben werden. Hier sollte die
%Beschreibung ``technischer'' bzw.~``formaler'' sein als in der
%Einleitung, da der Leser nun mit den Grundlagen und verwandten
%Arbeiten vertraut ist.

\section{Global Illumination Scheme}
\label{sec:scheme}
Within the scope of this work, we will develop a Global Illumination Scheme designed to efficiently propagate the intensities stored in discrete polar coordinates given as impulses (stimuli) by the user or the config. Polar profiles for different light sources are shown in Fig. \ref{polar}:
\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.8\textwidth]
    {img/strahler.png}
  }
  \caption{Polar Profiles for Different Types of Emitters}
  \label{polar}
\end{figure}\vskip 3pt
As a reminder, polar profiles map each angle $\varphi$ a magnitude $r$ forming a function $r(\varphi)$. They are circular axis versions of the domain $[0..2\pi]$, with negative magnitudes mirrored by 180 degrees). Thus, they are functions with an angular domain making them directionally dependent, which is a well suited representation for point lights, as we use them in a style of \cite{dachsbacher} et.al. For the sake of interpretability, we exclude negative values (energies) from the scope of our calculations. As a start-up for the propagation, we will think of a grid cell in 2D as a center pixel within an 8-neighborhood. It has 8 (4 faces/edges and 4 diagonals) unique neighbors, which are adjacent point lights in the grid (the type of light source with theoretically no finite extend). We consider one face neighbor (top) to shortly explain the steps of the scheme. The diagonals and the others are obtained from derivation and symmetry.

@@

%In diesem und den nachfolgenden Abschnitten werden die Beiträge der
%Arbeit motiviert, formal sauber (oft mathematisch, sprich mit
%Definitionen etc.) beschrieben, und bei Bedarf mithilfe von Beispielen
%verdeutlicht. Die Beschreibungen in diesem Kapitel sind meist
%unabhängig von einer konkreten Realisierung und Daten; diese werden im
%nachfolgenden Kapitel detailliert.

\section{Transmission Profiles and Weighting Functions}
\label{sec:trans}
To obtain a kind of footprint of the tensor field, we set up ellipse equations for each glyph representation, obtained from applying PCA on each cell (tensor) in the field. We map the singular values (corresponding to eigenvalues) in decreasing order to the ellipses radii $a(\lambda_1 )$ and $b(\lambda_2 )$ as follows:

Ellipse Equation:
\begin{align}
	r(\omega) = \frac{ab}{\sqrt{a^{2}\sin^{2}(\omega-\varphi)+b^{2}\cos^{2}(\omega-\varphi)}} =\frac{\lambda_1\lambda_2}{\sqrt{\lambda_1^2\sin^2(\omega-\varphi)+\lambda_2^2\cos^2(\omega-\varphi)}}
\end{align}
That is, the singular values form the half-axes of the PCA ellipsoid, which yields us a symbolic definition for our transmission (transfer) function as weighting profiles. We compute the offset angle $\varphi$ w.r.t. x-axis by exploiting the atan2-function with 4-quadrant evaluation for the first singular vector. As a preliminary prep, this profile is precomputed (sampled) for all discrete angular steps for each cell (e.g. $360$). The weighting scheme which we apply is depicted in Fig. \ref{weighting}.
\begin{figure}[h]
  \centering
  {
    \includegraphics[width=0.6\textwidth]
    {img/weighting.png}
  }
  \caption{red: Intensity Profile, cyan: Transmission Profile (Glyph Ellipsoid Eq.)}
  \label{weighting}
\end{figure}
We interpret the polar functions as discrete, sampled 1D vectors with a constant angular resolution. That means, we use element-wise multiplication for the 1D-vectors in red (Intensity Profile) and cyan (Transmission Profile) to obtain a windowed version of the intensity profile and perform the following integration step to accumulate the total radiant power (flux) as follows. Note that we model angular dependent radiant intensity ${I}(\omega) [\frac{W}{\mathop{rad}}]$ as an underlying physical entity, which is well suited for describing point light sources. The primary definition for our total accumulated incident radiant flux is denoted as:

\begin{align}
	\Phi_t &= \int_0^{2\pi}{I}(\omega){T}(\omega)\mathop{d\omega}\\
	\sum_{\cos} &= \int_0^{2\pi}\cos(\omega)\mathop{d\omega} \approx 2
\end{align}

This enables us to scale an oriented cosine-lobe w. the total radiant flux w. respect to its own total energy when integrated likewise. A discrete cosine lobe never reaches its theoretical sum of $2$ exactly in the interval $[0,2\pi]$. Therefore it needs to be computed to account for floating-point precision errors and normalized by following relation (e.g. $\sum_{\cos} \approx 1.9949204635834517$).
\begin{align}
	\cos_n(\omega) = \frac{\Phi_t}{\sum_{\cos}}cos(\omega)
\end{align}
This yields a cosine lobe, which has an overall energy of $\Phi_t$ if the point light is interpreted with a Lambert-Emitter (cnf. \cite{lindlein}) characteristic (as part or area element of a diffusely reflecting surface).

Now each tensor ellipse equation has a mean transmission rate $[\%]$, which turns out to be a measure or synonym for the total transmitted rate of the total incoming radiant intensity $\int_0^{2\pi}{I}(\omega)$. Since energy can not be generated from transmission profiles (only transmitting, not generating), we have to normalize every tensor ellipse equation with the highest mean transmission rate in the field. This restricts the tensor ellipse mean to a maximum of $100\%$ for the one with the highest tensor magnitude. We will use the definition of mean tensor magnitude $\mean{r}_{max}$ obtained from the PCA here for simplicity. Hence the tensor ellipse equations are normalized through the following step:
\begin{align}
	r_n(\omega) = \frac{1}{\mean{r}_{max}}r(\omega)
\end{align}
Notice, that this implements the concept of absorption implicitly, because tensors with a lower mean $\mean{r}$ than $100\%$ lose energy wheras absorption equals ${\mu} = {1}-{T}$ here. Now that we have coped with the problem of overall transmission, we are left with another problem that occurs with the weighting functions as that they can yield values above 1. This would potentially generate intensity out of nothing which leads to crucially unstable behaviour. Well, but only in case energy conservation principles were not respected.  To construct a solution, we ask for the source of the excessive energy. If the tensor mean complies with the constraint $\mean{r}\leq 1.0$, then the energy must lack in some other place (since the overall transmission fulfills the energy conservation constraint in that case). The only problem is that a normalization condition in usual style for the transmission profile ${T}(\omega)$ does not necessarily imply the same normalization for the term ${I}(\omega){T}(\omega)$. However, a proper normalization can give corrective remedy, wich reveals another subsequent normalization constraint. Hence the transmitted energy is given as:

\begin{align}
	{I}_t(\omega) &= n_f{I}(\omega){T}(\omega) \hskip 30pt w. \ n_f = \frac{\mean{T}\mean{I}}{\overline{TI}}
\end{align}
This transforms our equation for the radiant flux into:
\begin{align}
	\Phi_t &= \int_0^{2\pi}{I}_t(\omega){T}(\omega)\mathop{d\omega}
\end{align}
,which is now respecting energy conservation principles very conveniently as all we need is a set of means and we are done. The normalization factor is defined in 3 subsequent steps:
\begin{enumerate}
	\item Normalization of $TI$ to $\overline{TI}=1.0$
	\item Subsequent scaling w. mean intensity $\mean{I}$ for energy conservation principles
	\item Subsequent scaling w. mean tensor magnitude $\mean{T}$ for absorption principles
\end{enumerate}

This approach ensures that no energy will be generated and therefore implements the 	physical law of energy conservation or Kirchhoff's current law if you want to think of a particle's diffusion process. In fact the approach has proven to be robust w.r.t. energy losses for several iterations as depicted in chap. \ref{chap:eval} in the stochastic error graph.

\section{Global Illumination Gradient}
\label{sec:ftle}
Note that the resulting scalar field is $(n+1)$-dimensional ($x,y,z$ and direction $\theta$) and needs to be flattened by averaging or projection in 3D for proper visualization. 
Usw.
